QECO: A QoE-Oriented Computation Offloading Algorithm based on Deep Reinforcement Learning for Mobile Edge Computing

### 2024-11-16 Revision/Paper_Rev.tex
Rev_V8.6

\end{itemize}
The structure of this paper is as follows.  \textcolor{blue}{Section~\ref{section:II} reviews the related works.} Section~\ref{section:III} presents the system model, followed by the problem formulation in Section ~\ref{section:IV}. In Section~\ref{section:V}, we present the algorithm, while Section~\ref{section:VI} provides an evaluation of its performance. Finally, we conclude in Section~\ref{section:VII}. \color{blue}
\section{Related Work}
\label{section:II}
%\textcolor{black}{We explore state-of-the-art techniques characteristic to discuss how they address task offloading problems in scenarios that need to process massive data with limited devices in MEC, such as the Internet of Things (IoT), Internet of Vehicles (IoV), and Industrial IoT (IIoT)  environment. 
%Internet of Things (IoT) \cite{dai2020edge}, \cite{liao2023online}, \cite{Bolourian-WCL24}, 
%Internet of Vehicles (IoV) \cite{Bolourian-WCL24}, 
%and Industrial IoT (IIoT)  environment \cite{gong2022edge}, \cite{wu2023multi}, \cite{wu2024combining}
%, while uncertain requirements and sensitive latency properties of computation tasks pose nontrivial challenges in MEC systems with limited resources. 
% \textcolor{blue}{In recent years, edge computing has become a popular method for scenarios that need to process massive data with weak devices, such as IoTs \cite{zhang2023multi}, Internet of Vehicles (IoVs) \cite{lin2022multi}, \cite{wei2023many}, and Industrial IoT (IIoT) \cite{yuan2023adaptive} environment. We explore state-of-the-art RL-based techniques characteristic to discuss how they address task offloading problems in MEC. Table~\ref{table1} provides an intuitive comparison of the related works.}
  In recent years, numerous studies have focused on computation offloading in MEC, covering a broad range of applications, including IoTs \cite{zhang2023multi}, Internet of Vehicles (IoVs) \cite{lin2022multi}, \cite{wei2023many}, and Industrial IoT (IIoT) \cite{yuan2023adaptive}. 
%To take advantage of RL methods for computation offloading in MEC, most studies transfer the problem to MDP to model the interactions of the environment. 
To effectively tackle the challenges of MEC arising from the ever-changing nature of networks, recent research highlights the effectiveness of DRL in adapting to environmental changes and learning optimal strategies. In this section, we explore the current DRL-based state-of-the-art works and discuss their strengths and limitations. Table~\ref{table1} provides an intuitive comparison of these works.
\subsection{DRL-based Optimization Methods for MEC}

There are two primary categories of related research in resource management for MEC based on target objectives, single-objective optimization and multi-objective optimization methods, where RL are mainstream optimization techniques.
\subsubsection{Single-Objective Optimization} 
Recent research has extensively explored single-objective optimization in the context of MEC, focusing on optimizing a single objective, such as delay or energy consumption. 
Given the dynamic nature of MEC systems, in \cite{li2022integrated} the authors proposed an integrated optimization framework, which leverages past experience and model knowledge to enable fast and resilient real-time offloading control. 
%Li \textit{et al.} in \cite{li2020joint} introduced a DRL-based delay-optimal offloading strategy that employs a DQN to learn offloading decisions autonomously. 
To minimize total delay and reduce mobile vehicle task waiting time, Tang \textit{et al.} in \cite{tang2022double} developed a dynamic offloading model for multiple mobile vehicles, segmenting tasks into sequential subtasks for more precise offloading decisions.
Sun \textit{et al.} in \cite{sun2024hierarchical} explored both computation offloading and service caching problems in MEC. They formulated an optimization problem that aims to minimize the long-term average service delay. They then proposed a hierarchical DRL framework, which effectively handles both problems under heterogeneous resources.
To minimize energy consumption, Munir \textit{et al.} \cite{munir2021multi} developed a semi-distributed approach using a multi-agent RL framework for self-powered MEC. 
Zhou \textit{et al.} in \cite{zhou2021deep} proposed a Q-learning approach, which is an extension of RL to achieve optimal resource allocation strategies and computation offloading.
Wang et al. in \cite{wang2020reinforcement} studied the computation offloading problem in MEC systems within cellular networks, emphasizing mobile awareness. They proposed a joint approach for task offloading and computation allocation, aiming to address the task offloading costs associated with user mobility. 
In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility.
Dai \textit{et al.} in \cite{dai2020edge} introduced the integration of action refinement into DRL and designed an algorithm based on deep deterministic policy gradient (DDPG) to optimize resource allocation and computation offloading concurrently.

@@

\subsubsection{Multi-Objective Optimization.}
%In fact, the state of MD's computing and transmission power, arrived task's requirement, the energy state of MDs, and workloads at ENs all change dynamically with time in actual systems, 
%On the other hand, although the works [11]–[15] considered non-divisible tasks, they do not take into account the underlying queuing systems. As a result, the processing and transmission of each task should always be accomplished within one time slot (or before the arrival of the next task), which may not always be guaranteed in practice.
In MEC environments, multiple objectives often compete, requiring a balance in decision-making. For instance, task offloading must carefully navigate trade-offs between delay and energy consumption, as well as balance energy usage against the number of completed tasks. Research has increasingly addressed these multi-objective optimization challenges.
 Chen \textit{et al.} in \cite{chen2021drl} addressed a joint optimization problem involving computation offloading and resource allocation, aiming to reduce both task processing delay and energy consumption across all MDs.
To optimize privacy protection and quality of service, authors in \cite{wu2024privacy} investigated the joint computation offloading and power allocation problems for the IIoT network. They modeled the problem as an MDP and proposed a multi-agent DQN-based algorithm. 
 Huang \textit{et al.} in \cite{huang2019deep}, focused on a wireless-powered MEC and proposed a DRL-based approach, capable of attaining near-optimal decisions. This is achieved by selectively considering a compact subset of candidate actions in each iteration. 
 Liu \textit{et al.} in \cite{liu2021learn} investigated a two-timescale computing offloading and resource allocation problem and proposed a resource coordination algorithm based on multi-agent DRL, which can generate interactive information along with resource decisions. 
 Wu \textit{et al.} in \cite{wu2023computation} introduced a stochastic game-based resource allocation in the SDN-based MEC network. They used an MDP and proposed a multi-agent RL method to minimize both energy consumption and processing delay. 
 In \cite{Bolourian-WCL24}, the authors proposed an offloading algorithm using deep Q-learning for wireless-powered IoT devices in MEC systems. This algorithm aims to minimize the task drop rate while the devices solely rely on harvested energy for operation. 
%To reflecting real-world scenarios more accurately, the authors applied queuing theory to model MEC systems. It helps in modeling the dynamic nature of network, allowing MEC systems to allocate computational and transmission resources more efficiently.

To further enhance the modeling of real-world scenarios, some studies have applied queuing theory to model MEC systems, which helps capture the dynamic nature of the network and allows MEC systems to allocate computational and transmission resources more efficiently.
 Liao \textit{et al.} in \cite{liao2023online} introduced a double DQN (DDQN)-based algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency when minimizing both task computation delay and energy consumption. 
To address the privacy-aware computation offloading problem, Wu \textit{et al.} in \cite{wu2024combining} modeled the problem as an MDP and proposed a DQN-based method to optimize the computation rate and energy consumption in a queuing-based IIoT network.
Huang \textit{et al.} in \cite{huang2021deadline} proposed a DRL-based method based on a partially observable MDP (POMDP), which guarantees the deadlines of real-time tasks while minimizing the total energy consumption of MDs. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks. 
Tang \textit{et al.} in \cite{9253665} investigated the task offloading problem for indivisible and deadline-constrained computational tasks in MEC systems. The authors proposed a distributed DRL-based offloading algorithm designed to handle uncertain workload dynamics at the ENs. 
 Gong \textit{et al.} in \cite{gong2022edge} proposed a DRL-based network structure in the IIoT systems to jointly optimize task offloading and resource allocation to achieve lower energy consumption and decreased task delay.
 To optimize delay and energy consumption, Wu \textit{et al.} in \cite{wu2023multi} investigated the computation offloading problem in a queuing-based MEC IIoT system. They modeled the problem as a POMDP and proposed a multi-agent proximal policy optimization (PPO)-based method to obtain the optimal offloading strategy in dynamic environments.
 In \cite{gao2022large}, Gao \textit{et al.} introduced an attention-based multi-agent algorithm designed for decentralized computation offloading.

@@

 				\cite{li2020joint} & Caching and computation offloading & Minimize total system delay &  Multi-agent DQN & Not consider system energy consumtion \\ \midrule

@@

 				\cite{wang2020reinforcement} & Mobility aware computation offloading & Minimize energy consumtion & Q-learning & Not consider computation delay\\ \midrule

@@

 \subsubsection{Optimization Objectives}
Considering delay and energy consumption as optimization objectives represents a significant research focus in MEC. Most existing studies tend to optimize these two objectives independently, often neglecting the conflicts that arise between them, which can lead to skewed optimization results. Additionally, some research has prioritized maximizing the number of completed tasks, highlighting the advantages that MEC systems offer. Therefore, delay, energy consumption, and the number of completed tasks are critical factors in the design of MEC systems. However, to the best of our knowledge, there has been limited research that simultaneously addresses all three objectives.
Different from works \cite{li2022integrated}-\cite{dai2020edge}, which only focus on delay-optimal optimization or energy-efficient optimization, we are motivated to emphasize the MEC system with delay, energy consumption, and the number of completed tasks as three objectives for QoE optimization. Multi-objective studies \cite{chen2021drl}-\cite{gao2022large} have mostly investigated the trade-off between two or three of these factors but fail to properly address user demands and fulfill QoE requirements. QoE is a time-varying performance measure that reflects user satisfaction, where each user may have unique QoE requirements according to the Mobile Devices' (MDs) individual preferences.
In contrast to \cite{li2022integrated}-\cite{gao2022large}, to effectively address personalized QoE optimization, we define an adaptive trade-off between task completion and cost reduction. We introduce a weighted sum between completion delay and energy consumption as an associated cost. Each MD dynamically adjusts these weights to reflect the importance of different factors based on its energy modes. \color{black}

>>>>

\end{itemize}
The structure of this paper is as follows.  \textcolor{blue}{Section~\ref{section:II} reviews the related works.} Section~\ref{section:III} presents the system model, followed by the problem formulation in Section ~\ref{section:IV}. In Section~\ref{section:V}, we present the algorithm, while Section~\ref{section:VI} provides an evaluation of its performance. Finally, we conclude in Section~\ref{section:VII}. \color{blue}

\newpage.\newpage

%\textcolor{black}{We explore state-of-the-art techniques characteristic to discuss how they address task offloading problems in scenarios that need to process massive data with limited devices in MEC, such as the Internet of Things (IoT), Internet of Vehicles (IoV), and Industrial IoT (IIoT)  environment. 
	%Internet of Things (IoT) \cite{dai2020edge}, \cite{liao2023online}, \cite{Bolourian-WCL24}, 
	%Internet of Vehicles (IoV) \cite{Bolourian-WCL24}, 
	%and Industrial IoT (IIoT)  environment \cite{gong2022edge}, \cite{wu2023multi}, \cite{wu2024combining}
	%, while uncertain requirements and sensitive latency properties of computation tasks pose nontrivial challenges in MEC systems with limited resources. 
	% \textcolor{blue}{In recent years, edge computing has become a popular method for scenarios that need to process massive data with weak devices, such as IoTs \cite{zhang2023multi}, Internet of Vehicles (IoVs) \cite{lin2022multi}, \cite{wei2023many}, and Industrial IoT (IIoT) \cite{yuan2023adaptive} environment. We explore state-of-the-art RL-based techniques characteristic to discuss how they address task offloading problems in MEC. Table~\ref{table1} provides an intuitive comparison of the related works.}
	%To take advantage of RL methods for computation offloading in MEC, most studies transfer the problem to MDP to model the interactions of the environment.

	%\subsection{DRL-based Optimization Methods for MEC} 
	%\subsubsection{Single-Objective Optimization}

@@

	%\subsubsection{Multi-Objective Optimization.}

	%In fact, the state of MD's computing and transmission power, arrived task's requirement, the energy state of MDs, and workloads at ENs all change dynamically with time in actual systems, 
	%On the other hand, although the works [11]–[15] considered non-divisible tasks, they do not take into account the underlying queuing systems. As a result, the processing and transmission of each task should always be accomplished within one time slot (or before the arrival of the next task), which may not always be guaranteed in practice.

\color{black}
\section{Related Work}
\label{section:II}

%In recent years, numerous studies have focused on computation offloading in MEC, covering a broad range of applications, including IoTs \cite{zhang2023multi}, Internet of Vehicles (IoVs) \cite{wei2023many}, and Industrial IoT (IIoT) \cite{yuan2023adaptive}.

%Recent research has extensively explored single-objective optimization in the context of MEC, focusing on optimizing a single objective, such as delay or energy consumption.

%There are two primary categories of network performance optimization in MEC based on target objectives, single-objective optimization and conflicting multiple objective optimization approaches, where RL are mainstream optimization techniques.

To effectively tackle the challenges of MEC arising from the ever-changing nature of networks, recent research highlights the effectiveness of RL in adapting to environmental changes and learning optimal strategies. In this section, we explore the current RL-based state-of-the-art works and discuss their strengths and limitations. Table~\ref{table1} provides an intuitive comparison of these works.

Recent research has extensively explored single-objective optimization (SOO) in MEC, focusing on optimizing a single objective, such as delay or energy consumption.
Given the dynamic nature of MEC systems, in \cite{li2022integrated} the authors proposed an integrated optimization framework, which leverages past experience and model knowledge to enable fast and resilient real-time offloading control. 
%Li \textit{et al.} in \cite{li2020joint} introduced a DRL-based delay-optimal offloading strategy that employs a DQN to learn offloading decisions autonomously. 
To minimize total delay and reduce mobile vehicle task waiting time, Tang \textit{et al.} in \cite{tang2022double} developed a dynamic offloading model for multiple mobile vehicles, segmenting tasks into sequential subtasks for more precise offloading decisions.
Sun \textit{et al.} in \cite{sun2024hierarchical} explored both computation offloading and service caching problems in MEC. They formulated an optimization problem that aims to minimize the long-term average service delay. They then proposed a hierarchical DRL framework, which effectively handles both problems under heterogeneous resources.
To minimize energy consumption, Munir \textit{et al.} \cite{munir2021multi} developed a semi-distributed approach using a multi-agent RL framework for self-powered MEC. 
Zhou \textit{et al.} in \cite{zhou2021deep} proposed a Q-learning approach, which is an extension of RL to achieve optimal resource allocation strategies and computation offloading.
%Wang et al. in \cite{wang2020reinforcement} studied the computation offloading problem in MEC systems within cellular networks, emphasizing mobile awareness. They proposed a joint approach for task offloading and computation allocation, aiming to address the task offloading costs associated with user mobility. 
In \cite{zhao2019deep}, Zhao \textit{et al.} proposed a computation offloading algorithm based on DRL, which addresses the competition for wireless channels to optimize long-term downlink utility.
Dai \textit{et al.} in \cite{dai2020edge} introduced the integration of action refinement into DRL and designed an algorithm based on deep deterministic policy gradient (DDPG) to optimize resource allocation and computation offloading concurrently.

In contrast to these SOO studies, recent efforts have focused on conflicting multiple-objective optimization (CMOO) in MEC,  where RL are mainstream optimization techniques. CMOO often involves conflicting objectives, requiring careful balancing in decision-making. For example, task offloading must navigate trade-offs between delay and energy consumption, as well as balance energy usage with the number of completed tasks. 
Chen \textit{et al.} in \cite{chen2021drl} addressed a joint optimization problem involving computation offloading and resource allocation, aiming to reduce both task processing delay and energy consumption across all MDs.
To optimize privacy protection and quality of service, authors in \cite{wu2024privacy} investigated the joint computation offloading and power allocation problems for the IIoT network. They modeled the problem as an MDP and proposed a multi-agent DQN-based algorithm. 
 Huang \textit{et al.} in \cite{huang2019deep}, focused on a wireless-powered MEC and proposed a DRL-based approach, capable of attaining near-optimal decisions. This is achieved by selectively considering a compact subset of candidate actions in each iteration. 
 Liu \textit{et al.} in \cite{liu2021learn} investigated a two-timescale computing offloading and resource allocation problem and proposed a resource coordination algorithm based on multi-agent DRL, which can generate interactive information along with resource decisions. 
 Wu \textit{et al.} in \cite{wu2023computation} introduced a stochastic game-based resource allocation in the SDN-based MEC network. They used an MDP and proposed a multi-agent RL method to minimize both energy consumption and processing delay. 
 In \cite{Bolourian-WCL24}, the authors proposed an offloading algorithm using deep Q-learning for wireless-powered IoT devices in MEC systems. This algorithm aims to minimize the task drop rate while the devices solely rely on harvested energy for operation. 
%To reflecting real-world scenarios more accurately, the authors applied queuing theory to model MEC systems. It helps in modeling the dynamic nature of network, allowing MEC systems to allocate computational and transmission resources more efficiently.

To further enhance the modeling of real-world scenarios, some studies have applied queuing theory to model MEC systems, which helps capture the dynamic nature of the network and allows MEC systems to allocate computational and transmission resources more efficiently.
 Liao \textit{et al.} in \cite{liao2023online} introduced a double DQN (DDQN)-based algorithm for performing online computation offloading in MEC. This algorithm optimizes transmission power and scheduling of CPU frequency when minimizing both task computation delay and energy consumption. 
To address the privacy-aware computation offloading problem, Wu \textit{et al.} in \cite{wu2024combining} modeled the problem as an MDP and proposed a DQN-based method to optimize the computation rate and energy consumption in a queuing-based IIoT network.
Huang \textit{et al.} in \cite{huang2021deadline} proposed a DRL-based method based on a partially observable MDP (POMDP), which guarantees the deadlines of real-time tasks while minimizing the total energy consumption of MDs. This algorithm effectively tackles the challenges of dynamic resource allocation in large-scale heterogeneous networks. 
Tang \textit{et al.} in \cite{9253665} investigated the task offloading problem for indivisible and deadline-constrained computational tasks in MEC systems. The authors proposed a distributed DRL-based offloading algorithm designed to handle uncertain workload dynamics at the ENs. 
 Gong \textit{et al.} in \cite{gong2022edge} proposed a DRL-based network structure in the IIoT systems to jointly optimize task offloading and resource allocation to achieve lower energy consumption and decreased task delay.
 To optimize delay and energy consumption, Wu \textit{et al.} in \cite{wu2023multi} investigated the computation offloading problem in a queuing-based MEC IIoT system. They modeled the problem as a POMDP and proposed a multi-agent proximal policy optimization (PPO)-based method to obtain the optimal offloading strategy in dynamic environments.
 In \cite{gao2022large}, Gao \textit{et al.} introduced an attention-based multi-agent algorithm designed for decentralized computation offloading.

 In MEC environments, multiple objectives often compete, requiring a balance in decision-making. For instance, task offloading must carefully navigate trade-offs between delay and energy consumption, as well as balance energy usage against the number of completed tasks.

 Considering delay and energy consumption as optimization objectives represents a significant research focus in MEC.

 Most existing studies tend to optimize these two objectives independently, often neglecting the conflicts that arise between them, which can lead to skewed optimization results.

 Research has increasingly addressed these multi-objective optimization challenges.

@@

 				%\cite{li2020joint} & Caching and computation offloading & Minimize total system delay &  Multi-agent DQN & Not consider system energy consumtion \\ \midrule

@@

 				%\cite{wang2020reinforcement} & Mobility aware computation offloading & Minimize energy consumtion & Q-learning & Not consider computation delay\\ \midrule

@@

 \subsubsection{Optimization Objectives}
Considering delay and energy consumption as optimization objectives represents a significant research focus in MEC. Most existing studies tend to optimize these two objectives independently, often neglecting the conflicts that arise between them, which can lead to skewed optimization results. Additionally, some research has prioritized maximizing the number of completed tasks, highlighting the advantages that MEC systems offer. Therefore, delay, energy consumption, and the number of completed tasks are critical factors in the design of MEC systems. However, to the best of our knowledge, there has been limited research that simultaneously addresses all three objectives.
Different from works \cite{li2022integrated}-\cite{dai2020edge}, which only focus on delay-optimal optimization or energy-efficient optimization, we are motivated to emphasize the MEC system with delay, energy consumption, and the number of completed tasks as three objectives for QoE optimization. Multi-objective studies \cite{chen2021drl}-\cite{gao2022large} have mostly investigated the trade-off between two or three of these factors but fail to properly address user demands and fulfill QoE requirements. QoE is a time-varying performance measure that reflects user satisfaction, where each user may have unique QoE requirements according to the MDs' individual preferences.
In contrast to \cite{li2022integrated}-\cite{gao2022large}, to effectively address personalized QoE optimization, we define an adaptive trade-off between task completion and cost reduction. We introduce a weighted sum between completion delay and energy consumption as an associated cost. Each MD dynamically adjusts these weights to reflect the importance of different factors based on its energy modes. \color{black}
